---
title: "Assignment 1"
author: "Vaishnavi Behl"
student ID: "0304282847"
output: html_document
---

```{r setup, include=FALSE} 
# this chunk contains code that sets global options for the entire .Rmd. 
# we use include=FALSE to suppress it from the top of the document, but it will still appear in the appendix. 

knitr::opts_chunk$set(echo = FALSE) # actually set the global chunk options. 
# we set echo=FALSE to suppress code such that it by default does not appear throughout the document. 
# note: this is different from .Rmd default
```


# Exercise 2

**Analysing the built-in dataset "billboard" which contains song rankings for Billboard top 100 in the year 2000.**

In order analyse this dataset, we need to:

* load the tidyverse package and make it available in this session 
* load the billboard dataset

```{r}
#loading package
library(tidyverse)

#loading dataset
data("billboard")

#checking if data is tidy
head(billboard)
```

We can see that the above data is not tidy, since each row represents multiple observations - the artist, track, date and rank. 

In order to tidy the data, we need to create a separate coloumn for the variable "week" and another one for "rank". 

```{r}

#creating new coloumns 

billboard2 <- billboard %>% 
pivot_longer(cols=wk1:wk76, names_to = "week", values_to = "rank")

head(billboard2)
```


###Exercise 3

In this exercise, we're analyse the built in R dataset "trees" that provides measurements of the diameter, height and volume of timber in 31 felled black cherry trees.

```{r}

#Loading the dataset trees into memory
data("trees")

#analysing variables
head(trees)

```

```{r, echo=TRUE}

# Using piping operator to compare the height of the trees with the average girth and volume
trees %>%
  group_by(Height) %>%
  summarise(mean_Girth = mean(Girth),
            mean_Vol = mean(Volume))

head(trees)

# Calculating the height to volume ratio of trees between 70 to 80 feet long

trees %>%
  arrange(Girth) %>%
  mutate(Ratio = Height/Volume)

```

We can see that there is an inversely proportional relationship between the girth of black cherry trees and the ratio of their height to weight. 


## Appendix: All code in this assignment

```{r ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE} 
# this chunk generates the complete code appendix. 
# eval=FALSE tells R not to run (``evaluate'') the code here (it was already run before).
```
